%*******************************************************************************
%*********************************** First Chapter *****************************
%*******************************************************************************

\chapter{Plastic deformation}  %Title of the First Chapter
\label{chap:plastic_deformation}

\epigraph{I have ventured to call them dislocations}{A.E.H. Love}




\graphicspath{{plastic_deformation/Figs/}}

Plastic deformation or plasticity is the process of permanently altering the shape of a solid body under the influence of an applied external force. Indeed under the application of a large enough force and if cracking can be suppressed, for example by applying a confining pressure, most materials will plastically deform. Though a range of mechanisms for plastic deformation exist in crystals by far the most widespread is dislocation glide. Glide was first observed in 1867 \cite{Reusch1867} and was studied formally as early as 1899 \cite{Ewing1899,Ewing1900}. However it was not known or even proposed that dislocations, linear defects in a crystal structure, mediate plastic flow by moving over rational crystallographic planes \cite{Kelly2012ch7}.

If the process of dislocation glide cannot occur then a material is usually {brittle} and will fail by fracture or cracking, while if dislocation glide can occur a material is usually {ductile} and will fail by yielding. Brittleness vs ductility of a material is not a fixed property but will depend on the stress state and the temperature. If sufficiently high hydrostatic pressure is applied then even materials like sapphire will undergo plastic flow \cite{Bridgman1947}. At high temperatures, thermal activation can enable glide to occur in materials that are brittle at lower temperatures, with many materials known to exhibit ductile to brittle transition temperatures, for example body-centre iron, titanium carbide and olivine among many others   \cite{Kelly2012ch7,frost1982,Rowcliffe1971,Darot1985}.

Dislocation motion is of much practical importance in crystalline materials. In many metals, particularly those with a face-centred cubic structure such as copper or aluminium, dislocation motion must be hindered to achieve sufficient strength for use as an engineering material. Much of physical metallurgy is the study of the microstructural features that affect dislocation motion and their genesis in materials processing or alloy composition.

In other materials the crystal structure itself provides such a large barrier to motion that almost no plastic deformation is possible and these materials usually fail by brittle fracture. This is true of many non-metallic materials widely used as protective coatings or as functional materials in devices. Fracture is often the life limiting factor for these materials, thus if their toughness were increased by making plastic flow were easier their lifetime might be extended. 

Even in some metallic materials with comparatively simple crystal structures plastic flow is limited not by microstructural features but by the inherent resistance of the crystal structure to the motion of dislocations. This is common in body-centred cubic metals, tantalum and niobium at room temperature and iron at lower temperatures \cite{Christian1983,Weinberger2013}, but also in the face-centred cubic metal iridium; the low mobility of dislocations in iridium is at least partly responsible for its tendency to fail by cleavage \cite{Panfilov2001}. The ductile to brittle transition in many metals is due to the lattice resistance, notably iron in its body-centred form becomes brittle at low temperatures. This resulted in the failure of the Liberty ships during the second world war in the cold waters of the North Atlantic. Originally thought to be due to high stresses caused by the welding technique used to join the steel, it was Constance Tipper who showed that it was the lack of plastic flow around the crack tip that enabled for the catastrophic failure of entire ships to occur \cite{Cottrell1997}. 

Thus there is a great motivation to understanding the inherent resistance to dislocation flow in a crystal structure, or \emph{lattice resistance}, and in being able to alter that lattice resistance and thereby introduce a degree of toughness to otherwise brittle phases.



\section{Dislocations}  
\label{sec:dislocations}
\input{plastic_deformation/dislocations.tex}


\section{Ductility criteria}
\label{sec:ductility_criteria}

\input{plastic_deformation/ductility_criteria.tex}



\section{Tailoring the Peierls stress}
\label{sec:tailor_peierls}
\input{plastic_deformation/tailoring_peierls.tex}



\section{Layered crystals}
\label{sec:layered_crystals}
\input{plastic_deformation/layered_crystals.tex}


\section{Alternative energy formulations}

\label{sec:empirical_potentials}
\input{plastic_deformation/alternative_energy.tex}

























































































