% ************************** Thesis Acknowledgements **************************
\clearpage

\doublespacing

\section*{Acknowledgements}


Firstly I'd like to thank Bill for his enthusiasm and advice. His approach to science in particular and life in general has made the research group a great place to be. Essential to my project were his willingness to return a problem to basics and patience in helping me think hard about crystals. Without this approach the problems would surely have spiralled out of control.


I'm grateful to all the people that made the Gordon Lab and the department the great place it is: John, Claire, Letitia, Rob, Philip and Joe who helped me settle in way back when, and those of you still here: Tom, Sam, Jimmy, Fernando, Alastair, Megan, Noel, Nashid, Eleanor, Max, Matt, Julia and everyone else. 
I'd particularly like to thank Joseph Reed and Thomas Furnival for years of friendship and ``useful discussions'', usually with \emph{just one} drink on the way home.

Thanks are also owed to my other friends who've put up with me through all this, Adam Greig, Max Conway and Laurie Burchell particularly, and to all my other friends who I'm so lucky to know.

Finally I'd like to say how grateful I am to my family for all the support over the years.
Mum and Dad, you've helped me to be curious about the world and showed me that we can describe and understand it. You set me on the road to being a scientist, but also kept me sane and happy over this near decade of academia, providing a place to take refuge when I found I still needed to be looked after. I'd also like to thank my grandparents who have all been so important to me, Peter and Barbara, and Alan and Marie.


Thank you all.

























































